---
title: "5234 project"
author: "Yiran Su"
date: "2023-04-11"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(purrr)
```

#Import dataset
```{r}
ipums<- read.csv('/Users/susu/Desktop/final projects/5234/ipums.csv')
dim(ipums); head(ipums)
```



#PART ONE: CLUSTER SAMPLING
#(1A) Generate a frequency table of the number of persons within each psu.
```{r}
freq <- table(ipums$psu)
freq
sum(freq)
```

#(1B) Take a two-stage cluster sample of size 200
```{r}
set.seed(1234)
n=10; (N=nrow(freq))
(sample <- sample(1:N, n,replace = FALSE))
psu_sample <- ipums[ipums$psu %in% sample,]

M0<-nrow(psu_sample)

#data is the column of nested data frames for each PSUs in ssu_sample.
#sample_n is the function that is applied to each nested data frame in data to sample mi observations without replacement.

ssu_sample <- psu_sample%>%
  group_by(psu)%>%
  nest()%>%
 mutate(Mi	=	map_dbl(data,	nrow),
        pi = Mi / M0,
        mi= round(200*pi))%>%
  mutate(second_stage= map2(data, mi, sample_n, replace = FALSE))%>%
  select(-data) %>%
  unnest(second_stage)

nrow(ssu_sample)

```


#(1C)estimate the population mean and total of ‘inctot’ and give the standard errors of your estimates
```{r}
n=10; N=nrow(freq)
(ybar.i=aggregate(inctot~psu,ssu_sample,mean)$inctot)

(t.i=aggregate(inctot~psu,ssu_sample,sum)$inctot)
(M.i=aggregate(Mi~psu,ssu_sample,mean)$Mi)
s.i2=aggregate(inctot~psu,ssu_sample,sum)$inctot
(that.i=M.i*ybar.i)
(that.unb=N/n*sum(that.i))#estimated population total
s.t2=sum((that.i-that.unb/N)^2)/(n-1)
(se.t.hat.unb=sqrt(N/n*sum((1-n/M.i)*(s.i2/n)*M.i^2)+ N^2*(1-n/N)*s.t2/n))#standard error of population total income estimator


(ybar.hat.r=sum(that.i)/sum(M.i)) #estimated population mean
s.r2=sum((that.i-M.i*ybar.hat.r)^2)/(n-1)
Mbar=sum(M.i)/n
b=sum((1-n/M.i)*M.i^2*s.i2/n)
var.ybar.hat.r=1/Mbar^2*((1-n/N)*s.r2/n+1/N*b/n)
(se.ybar.hat.r= sqrt(var.ybar.hat.r)) #standard error of population average income estimator 
```

#(1D) Select an unequal-probability sample of 10 psus, with probability proportional to number of persons within each psu, and then take a subsample (SRS) of 20 persons in each of the selected psus. 


```{r}
set.seed(1)
psaii=freq/sum(freq)
(unequal_sample<- sample(1:90,10,replace= FALSE,prob=psaii))
unequal_psu_sample <- ipums[ipums$psu %in% unequal_sample,]

M0<-nrow(psu_sample)

#data is the column of nested data frames for each PSUs in ssu_sample.
#sample_n is the function that is applied to each nested data frame in data to sample mi observations without replacement.

unequal_ssu_sample <- unequal_psu_sample%>%
  group_by(psu)%>%
  nest()%>%
 mutate(Mi	=	map_dbl(data,	nrow),
        pi = Mi / M0,
        mi= round(200*pi))%>%
  mutate(second_stage= map2(data, mi, sample_n, replace = FALSE))%>%
  select(-data) %>%
  unnest(second_stage)

nrow(unequal_ssu_sample)
```


#(E) Estimate the population mean and total of “inctot” with se
```{r}
n=10; N=nrow(freq)
(ybar.i.unequal=aggregate(inctot~psu,unequal_ssu_sample,mean)$inctot)

(t.i.unequal=aggregate(inctot~psu,unequal_ssu_sample,sum)$inctot)
(M.i.unequal=aggregate(Mi~psu,unequal_ssu_sample,mean)$Mi)
s.i2.unequal=aggregate(inctot~psu,unequal_ssu_sample,sum)$inctot
(that.i.unequal=M.i.unequal*ybar.i.unequal)
(that.unb.unequal=N/n*sum(that.i.unequal))#estimated population total
s.t2.unequal=sum((that.i.unequal-that.unb.unequal/N)^2)/(n-1)
(se.t.hat.unb.unequal=sqrt(N/n*sum((1-n/M.i.unequal)*(s.i2.unequal/n)*M.i.unequal^2)+ N^2*(1-n/N)*s.t2.unequal/n))#standard error of population total income estimator


(ybar.hat.r.unequal=sum(that.i.unequal)/sum(M.i.unequal)) #estimated population mean
s.r2.unequal=sum((that.i.unequal-M.i.unequal*ybar.hat.r.unequal)^2)/(n-1)
Mbar.unequal=sum(M.i.unequal)/n
b.unequal=sum((1-n/M.i.unequal)*M.i.unequal^2*s.i2.unequal/n)
var.ybar.hat.r.unequal=1/Mbar.unequal^2*((1-n/N)*s.r2.unequal/n+1/N*b.unequal/n)
(se.ybar.hat.r.unequal= sqrt(var.ybar.hat.r.unequal)) #standard error of population average income estimator 

```



#PART TWO: COMPLEX SAMPLING
#(2A) Select a two-stage stratified cluster sample as follows: first select two psus from each stratum, with probability proportional to size, and then take a simple random sample of 50 persons from each selected psu. 

```{r}
library(ggplot2)
library(dplyr)
library(survey)



```


```{r}
```


#(2B) Construct the column of sampling weights for your sample.
```{r}

```


#(2C) Draw a histogram of ‘inctot’ for your sample that incorporates weights.
```{r}

```

#(2D) Draw two of the scatterplots that incorporate weights for 𝑦𝑦 variable ‘inctot’ and 𝑥𝑥 variable ‘age’. How do these differ from scatterplots that do not use the weights?
```{r}

```

#(2F) Create a categorical variable catinc from ‘inctot’ with two categories: low income and high income, using the median income as the dividing point for the categories.
```{r}

```


#(2G) Using your sample, conduct a hypothesis test to explore whether ‘catinc’ is associated with ‘race’. What method did you use to account for the complex sampling design?
```{r}

```

